### 2020-10-27 submission/manuscript.tex

\documentclass[]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
% Redefine \includegraphics so that, unless explicit options are
% given, the image width will not exceed the width of the page.
% Images get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\ScaleIfNeeded{%
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother
\let\Oldincludegraphics\includegraphics
{%
 \catcode`\@=11\relax%
 \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
}%
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

\author{}
\date{\vspace{-2.5em}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\linenumbers
\modulolinenumbers
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

\begin{document}

\section{Removal of rare sequences from 16S rRNA gene sequence surveys
biases the interpretation of community structure
data}\label{removal-of-rare-sequences-from-16s-rrna-gene-sequence-surveys-biases-the-interpretation-of-community-structure-data}

\vspace{35mm}

Patrick D. Schloss${^\dagger}$

\vspace{40mm}

$\dagger$ To whom correspondence should be addressed:

\href{mailto:pschloss@umich.edu}{pschloss@umich.edu}

Department of Microbiology and Immunology University of Michigan Ann
Arbor, MI 48109

\vspace{35mm}

\subsection{Research article format}\label{research-article-format}

\newpage
\linenumbers

\subsection{Abstract}\label{abstract}

\textbf{word choice: sample vs.~community}

250 words

\newpage

\subsection{Importance}\label{importance}

150 words

\newpage

\textbf{EDIT: Two elements that are held in tension in the analysis of
16S rRNA gene sequence data is how to adequately remove PCR and
sequencing artifacts and decrease the granularity of the taxonomic level
that is used in the analysis.} Previous attempts have included screening
for sequencing quality based on quality scores {[}Kozich/Edgar{]}
followed by a polishing step based on the frequency of the sequences
relative to similar sequences {[}Kozich/Edgar/DeBlur{]}. Other pipelines
model the quality scores and types of errors to cluster sequences
directly {[}Dada2{]}. But, as a final step many pipelines advocate for
removing rare sequences from each dataset prior to outputting the
sequence data as amplicon sequence variants (ASVs)
{[}Knight/Edgar/DeBlur/Dada2{]}. ASVs are often clustered further to
generate operational taxonomic units or phylotypes. Some pipelines
remove all ASVs that appear once (i.e.~singletons) {[}XXXX{]}, XXXXXX
{[}XXXXXX{]}, XXXXX {[}XXXXXXX{]}, or XXXXXX {[}XXXXXX{]} times prior to
further clustering or making ecological comparisions. Notably, the
mothur-based pipeline discourages the practice of removing rare
sequences.

The abundance-based screening approach assumes that rare ASVs are more
likely to be artifacts than more abundant ASVs. Sequencing of mock
communities confirms that artifacts tend to be rare. Proponents of
abundance-based screening point to their ability to obtain the correct
number of ASVs, OTUs, or phylotypes with data generated from sequencing
mock communities when rare ASVs are removed. However, this approach
effectively overfits the curation pipeline to data generated from a
phylogenetically simple community with an atypical community
distribution that is often sequenced to a depth that is not achieved
with biological samples. It is necessary to think more deeply about the
practice of abundance-based screening.

The minimum abundance thresholds that have been proscribed are developed
and applied without regard for the total number of sequences generated
from each sample. Ignored in their recommendations is the common
experience that the number of sequences generated from each sample may
vary by two or three orders of magnitude. An ASV that appears once in a
sample with 2,000 sequences is more trustworthy than an ASV that appears
once in a sample with 100,000 sequences since it has a 50-fold higher
relative abundance. But, according to the pipeline recommendations, they
are treated as being equally trustworthy. Rather than removing rare
ASVs, the approach taken by the mothur pipeline applies the classical
ecological approach of rarefaction. Each sample is rarefied to the same
sequencing depth so that the number of artifacts that appears in each
sample is controlled.

Experience sequencing biological samples demonstrates that there are
good ASVs that may have an abundance below the proscribed threshold. For
example, the abundance of an ASV may be below the threshold in some
samples or time points and above the threshold in others. However,
rarity, both in terms of prevalence and incidence, is an important
ecological concept. Removing rare ASVs likely hinders one's ability to
ability to make inferences about the dynamics and nature of the
populations that rare ASVs represent. Furthermore, removing ASVs whose
abundances are below the proscribed threshold also potentially biases
the community structure of the samples.

In the current study, I use published sequence data from 12 studies to
investigate the nature of rare ASVs (i.e.~those that appear 10 or fewer
times) and the effects of removing them on downstream analysis of
microbial communities. The analysis was also performed using traditional
operational taxonomic units, where ASVs subjected to abundance-based
screening were clustered such that the ASVs within an OTU were no more
than 3\% different from each other. The results reject the assumptions
built into abundance-based screening and highlight the problems inherent
in removing rare ASVs.

\subsection{Results}\label{results}

\textbf{Datasets.} I collected 12 publicly available datasets that used
the Illumina MiSeq platform to sequence the V4 region of the 16S rRNA
gene from a variety of environments (Table 1). To insure the highest
possible data quality, datasets were limited to those where the 500
cycle v2 MiSeq chemistry was used to sequence the amplicons. The paired
250 nt reads resulted in near complete 2-fold sequencing coverage of
every nucleotide in the ca. 250 nt-long region. This region and
sequencing platform were selected because previous work has shown that a
standard data analysis pipeline in mothur results in a sequencing error
rate below 0.02\% (1). All sequence data were obtained from the Sequence
Read Archive and processed using a standard mothur-based sequencing
pipeline that resulted in ASVs as generated by the pre.cluster algorithm
(1, 2). After removing poor quality and chimeric ASVs and samples that
had uncharacteristically low number of sequences for the dataset, these
datasets included between 7 and 490 samples (Figure S1). The median
number of sequences for each dataset ranged between 6,477 and 193,464
(Table 1). Strikingly, aside from the relatively small marine and soil
datasets, the difference between the sample with the fewest sequences
and the sample with the most sequences for each dataset varied by
between 7.4 and 96.6-fold (Table 1).

\textbf{The nature of singletons.} Removal of rare ASVs is commonly
justified as a method of removing ASVs that are artifacts. If such ASVs
are artifacts, then one would expect the number of singleton ASVs to
accumulate with sequencing depth. Contrary to this expectation, the
median percentage of sequences that were discarded when singleton ASVs
were removed from each dataset varied between 0.42 and 22.23\%
(bioethanol and seagrass). In addition, with the exception of the
samples from the marine and sediment datasets (Spearman correlation,
P\textgreater{}0.05), the fraction of singleton ASVs in samples was
negatively correlated with the number of sequences in each sample with a
range between -0.27 and -0.87 (rice and bioethanol) (Figure 1A). This
showed that with additional sequencing, the probability of seeing
singleton ASVs in multiple samples was greater than the probability of
generating an artifact. This suggests that the singleton ASVs are not as
likely to be artifacts as previously thought. Furthermore, if singleton
ASVs were artifacts, then one would not expect to find them in other
samples from the same dataset. In fact, singleton ASVs from samples with
fewer sequences were often found in samples with more sequences. At
least 50\% of the singleton ASVs found in the samples from the mice,
rice, seagrass, and stream datasets were found in another sample from
the same dataset (Figure 1B). Considering the likelihood of finding an
ASV duplicated in another sample is confounded by the number of samples
and inter-sample diversity, the high coverage of singleton ASVs in these
datasets was remarkable. The correlation between the number of sequences
in a sample and the fraction of that sample's singleton ASVs that were
covered by another sample in the dataset was significant and negative
for 9 of the datasets ranging between -0.31 and -0.84 for the rice and
seagrass datasets, respectively (Figure 1C). The negative correlation
indicated that the singleton ASVs in the smaller samples were more
likely to be covered by ASVs in the larger samples. Among the three
datasets without a significant correlation (Spearman correlation,
P\textgreater{}0.05), the marine and soil datasets had the fewest
samples in our collection and the stream dataset already had a high
level of coverage regardless of the number of sequences. Contrary to the
common motivation for removing rare ASVs, these results indicate that
this practice disproportionately impacts samples with fewer sequences
and likely removes more non-artifact ASVs than those that are artifacts.

\textbf{The impact of removing rare ASVs on the information represented
in each sample.} Removing rare ASVs will reduce the richness of ASVs
(i.e.~the number of ASVs per sample) and increase the relative abundance
of the remaining ASVs. To quantify the effect of removing rare ASVs on
the information contained within each sample, I varied the minimum
abundance threshold to simulate removing ASVs of varying rarity from
each sample. The richness of ASVs in each sample decreased by between
34.4 and 86.2\% (peromyscus and soil) when removing those ASVs that only
appeared once and by between 76.0 and 95.6\% (sediment and soil) when
removing those that appeared ten or fewer times from each sample (Figure
2A). Similarly, the Shannon diversity decreased by between 1.8 and
15.9\% (human and soil) when removing ASVs that only appeared once and
by between 5.4 and 35.4\% (human and seagrass) when removing ASVs that
appeared ten or fewer times from each sample (Figure 2B). Next, I
assigned the ASVs to OTUs, which were defined as a group of ASVs that
were more than 97\% similar to each other to assess the impact of
removing rare ASVs on higher level taxonomic groupings that are commonly
used in microbial ecology studies. Although pooling similar ASVs into
OTUs reduced the impact of removing the rare ASVs relative to the
ASV-based analysis, the minimum abundance threshold still decreased the
richness of OTUs and the diversity decreased relative to the full
community (Figure S2AB). In contrast to the richness and diversity
measurements, the Kullback--Leibler divergence compares the relative
abundance of specific ASVs or OTUs between representations of the
community. I calculated the Kullback--Leibler divergence between the
full communities and those where rare ASVs were removed. As the
threshold for removing ASVs increased, the amount of information lost
also increased for both ASVs and OTUs (Figure 2C and Figure S2C). The
relative loss of information was generally smaller for OTUs than than it
was for ASVs. Removing rare ASVs, regardless of abundance threshold, had
profound impacts on the representation of the communities.

\textbf{Removing treatment group effects from community data.} Because
treatment effects often affect a sample's diversity and inter-sample
variation, I generated null distributions for each study by randomizing,
without replacement, the number of times each ASV was observed in each
sample such that the total number of sequences in each sample and the
total number of times each ASV was observed across all samples in the
study was the same as was originally observed. This effectively made
every community in a study a statistical sample of the study-wide
composite community distribution. For example, after this procedure, the
490 samples from the human dataset would be expected to have the same
richness and diversity of ASVs and one would not expect to find
treatment-based effects between the samples. Because of the risk of bias
if only one representation of the null distribution was generated, I
generated 100 randomized datasets for each study. The trends between
removing rare ASVs and the richness, diversity, and information loss
that were identified using the observed community community distribution
data were also identified with the data from the null distribution;
however, the losses were larger when using the null distribution data
(Figure S3). The null distribution data were used in the remainder of
the study to minimize the risk of bias.

\textbf{The impact of removing rare ASVs on the information represented
between samples.} Considering the loss of richness, diversity, and
information when a community has its rarest ASVs removed, it seemed
likely that the relationship between communities would also be altered.
To assess the impact of removing rare ASVs on measures of alpha
diversity between samples I calculated the coefficients of variation
(COVs, i.e.~the standard deviation divided by the mean) for richness and
diversity for each study at multiple abundance thresholds. The COVs for
the richness of ASVs across the studies after removing singletons were
between 3.6 and 32.7-times larger than they were without removing
singleton ASVs (mice and stream; Figure 3A). Similarly, the COVs for the
diversity of ASVs were between 1.8 and 20.4-times larger when singletons
were removed than when they were not removed (mice and rice; Figure 3B).
To assess the impact of removing rare ASVs on measures of beta diversity
between samples, I calculated the COVs of the Bray-Curtis distances
between samples within the same study at multiple abundance thresholds.
The COVs between Bray-Curtis distances within a study when singletons
were removed was between 1.3 and 18.6-times larger than when they were
not removed (mice and stream; Figure 3C). When ASVs were clustered into
OTUs the difference in COVs was less than it was for the ASVs (Figure
S4). These results indicate that removing rare ASVs increases the
dissimilarity between samples, which could have a significant impact on
the statistical power to detect differences between treatment groups.

\textbf{The impact of removing rare ASVs on the ability to detect
statistically significant differences between treatment groups.} To test
the effect of increased inter-sample variation, I randomly assigned
samples to one of two treatment groups. In the first treatment group,
communities were randomly sampled from the null distribution as
described above. For the second treatment group, I randomly selected
10\% of the ASVs in the pooled study distribution to increase their
abundance by 5\%. I randomly generated 100 simulated sets of treatment
groups and samples. I then tested the ability to detect a difference
between the two treatment groups using alpha and beta diversity metrics.
The fraction of significant tests was a measurement of the statistical
power to detect the difference between the treatment groups. When
considering the differences in richness and diversity, the marine
dataset yielded no simulated sets that were statistically significant,
which was likely due to the small number of samples in the study (N=7).
Among the remaining datasets, the power to detect a difference in the
richness of ASVs ranged between 0.10 and 0.49 (sediment and stream) and
between 0.10 and 0.53 (rainforest and stream) to detect a difference in
diversity when using a Wilcox test (Figure 4A). When singleton ASVs were
removed, the power to detect a difference in the diversity of ASVs
dropped by between 27.3 and 92.9\% (bioethanol and soil) and by between
40.0 and 93.3\% (rainforest and soil; Figure 4B). The effect of removing
rare ASVs on the richness of OTUs and their diversity was similar
(Figure S5AB). I used the Bray-Curtis dissimilarity index to compare the
simulated communities within each dataset and calculated the power to
detect differences between the two simulated treatment groups using the
analysis of molecular variance (also called PERMANOVA) (Figure 4C and
S5C). Without removing rare sequences, the power to detect a difference
between the two simulated treatment groups varied between 0.41 and 1.00
(rainforest and rainforest). Aside from the bioethanol, human, and mice
datasets, the power to detect differences dropped by between 6.5 and
64.0\% (soil and rice) when singletons were removed. However, when ASVs
that occurred 10 or fewer times were removed from each sample, the power
to detect differences dropped by 12.0 and 97.2\% (human and peromyscus);
similar results were observed when ASVs were clustered into OTUs.
Removing rare ASVs reduced the ability to detect simulated treatment
effects using metrics commonly used to compare microbial communities.

\textbf{The impact of removing rare ASVs on the probability of falsely
detecting a difference between treatment groups.} I next asked whether
removing rare ASVs could lead to falsely claiming that a treatment
effect had a significant effect on community diversity and structure.
First, I sampled sequences from the null distribution for each dataset
and randomly assigned each sample to one of two treatment groups and
determined the richness and diversity of ASVs and OTUs. Testing at an
experiment-wise error rate of 0.05, I expected 5\% of the iterations for
each dataset to yield a significant test result. Indeed, there was no
evidence that removing rare ASVs resulted in an inflated experiment-wise
error rate. The average fraction of significant tests did not
meaningfully vary from 0.05 across the minimum abundance threshold,
dataset, metric of describing sample alpha-diversity, or whether the
abundance of ASVs or OTUs were used (Figure 5A and S6A). Similarly, the
average fraction of significant tests did not meaningfully vary from
0.05 when using analysis of molecular variance to compare communities
using Bray-Curtis distances (Figure 5A and S6A). Second, I again sampled
sequences from the null distribution, but assigned samples to one of two
treatment groups based on the number of sequences in each sample. The
samples with fewer than the median number of sequences for the dataset
were assigned to one group and those with more than the median were
assigned to the other. This exaggerated bias has been observed in
comparisons of the lung and oral microbiota because of the larger number
of non-specific amplicons that can be sequenced from lung samples
relative to those in the oral cavity leading to a significant difference
in sequencing depth between treatment groups {[}{]}. When rare sequences
were not removed, the fraction of significant tests did not differ from
5\% for comparing the richness, their diversity, or Bray-Curtis
distances (Figure 5B and S6B). However, when rare taxa of any frequency
were removed, the probability of falsely detecing a difference as
signifiant increased with the definition of rarity (Figure 5B and S6B).
Not including the small marine dataset, the average fraction the average
fraction of falsely detecting a difference across datasets when only
singletons were removed was 92.45\%. If there is any relationship
between the number of sequences and the treatment group, the risk of
falsely rejecting the null hypothesis is inflated when researchers use
the strategy of removing rare sequences. The most conservative approach
is to not remove low abundance sequences.

\subsection{Discussion}\label{discussion}

Removing rare sequences decreases the diversity represented by 16S rRNA
gene sequence data and increases the variation between samples. Such
impacts will hinder the statistical power to differentiate between
treatment groups. Instead of removing rare sequences, researchers should
focus on optimizing their sequence generation to minimize the amount of
PCR and sequencing errors. In addition, samples should be rarefied to a
common number of sequences across samples without prior culling of rare
sequences. The number of artifacts is correlated to the number of
sequences being considered. With this in mind, rarefaction allows one to
control for uneven sampling effort and to control for the number of
artifacts in the analysis.

Need to treat rare sequences with a grain of salt

\newpage

\subsection{Materials and Methods}\label{materials-and-methods}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  sequencing pipeline description
\end{itemize}

\subsection{Acknowledgements}\label{acknowledgements}

\newpage

\textbf{Table 1. Summary of studies used in the analysis.} For all
studies, the number of sequences used from each study was rarefied to
the smallest sample size. A graphical represenation of the distribution
of sample sizes for each study and the samples that were removed from
each study are provided in Figure S1.

\begin{tabular}{lrrrrr}
\toprule
\textbf{Study} & \textbf{Samples} & \textbf{\makecell[c]{Total\\sequences}} & \textbf{\makecell[c]{Median\\sequences}} & \textbf{\makecell[c]{Range of\\sequences}} & \textbf{\makecell[c]{Fold-difference\\between largest\\and smallest sample}}\\
\midrule
Bioethanol & 95 & 3,972,943 & 16,015 & 3,688-356,136 & 96.6\\
Human & 490 & 20,909,768 & 32,505 & 10,523-430,415 & 40.9\\
Lake & 52 & 3,169,868 & 69,041 & 15,347-112,871 & 7.4\\
Marine & 7 & 1,391,396 & 193,464 & 133,516-254,060 & 1.9\\
Mice & 348 & 2,813,747 & 6,477 & 1,804-30,565 & 16.9\\
Peromyscus & 111 & 1,555,545 & 12,446 & 4,464-33,644 & 7.5\\
Rainforest & 69 & 946,295 & 11,561 & 4,932-37,767 & 7.7\\
Rice & 490 & 22,591,168 & 43,216 & 2,776-193,464 & 69.7\\
Seagrass & 286 & 4,130,454 & 13,567 & 1,803-45,191 & 25.1\\
Sediment & 58 & 1,154,174 & 17,584 & 7,685-68,321 & 8.9\\
Soil & 18 & 956,656 & 51,844 & 47,806-59,956 & 1.3\\
Stream & 201 & 21,162,574 & 90,159 & 9,175-390,964 & 42.6\\
\bottomrule
\end{tabular}

\newpage

\textbf{Figure 1.}

\newpage

\textbf{Figure 2.}

\newpage

\subsection{References}\label{references}

1. \textbf{Kozich JJ}, \textbf{Westcott SL}, \textbf{Baxter NT},
\textbf{Highlander SK}, \textbf{Schloss PD}. 2013. Development of a
dual-index sequencing strategy and curation pipeline for analyzing
amplicon sequence data on the MiSeq Illumina sequencing platform.
Applied and environmental microbiology \textbf{79}:5112--5120.

2. \textbf{Schloss PD}, \textbf{Westcott SL}, \textbf{Ryabin T},
\textbf{Hall JR}, \textbf{Hartmann M}, \textbf{Hollister EB},
\textbf{Lesniewski RA}, \textbf{Oakley BB}, \textbf{Parks DH},
\textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres B},
\textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{http://dx.doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.

\end{document}